Efficient Calculation of Polynomial and Interaction Features on Sparse Matrices

### 2015-09-21 paper/paper2.tex
double dolumn
added pagenum to fig reference
back to one col

\documentclass[11pt]{article}

\usepackage{array}
\usepackage{clrscode3e}
\usepackage{amsmath}
\usepackage{kbordermatrix}
\usepackage{mathtools}
\usepackage{cite}
\usepackage{hyperref}
\usepackage{rotating}
\usepackage{svg}
\usepackage{pdfpages}

@@

\section{Introduction}
When performing a modeling task, one is often faced with a non-linearly separable problem,
or is attempting to predict a target that is a non-linear combination of the variables at one's disposal.
There are several general classes of methods for dealing with these situations: kernel methods, 
high variance models, and input transformations.

@@

\begin{sidewaysfigure}
    \centering
    \includegraphics[scale=0.5]{classifier_boundaries.png}
    \caption{Decision boundaries for four classifier types on four different datasets. The darker the color, the more certain the decision.
             The plain dataset is shown in the first column. Each other column is a type of classifier applied to that dataset. The classifier
             types are logistic regression, logistic regression with second degree polynomial features, RBF SVM, and gradient boosting with decision trees.
             The accuracy of each model is shown in the bottom right corner of that model's subplot.}
    \label{fig:boundaries}
\end{sidewaysfigure}

@@

@@

\newpage
INSERT JOHN'S PROOF HERE

@@

\newpage

@@

\subsection{Empirical}
To assess the performance of the algorithm on data of varied density, we compare its speed
to the PolynomialFeatures class of the popular machine learning scikit-learn \cite{scikit-learn}.
Uniformly random matrices of size $1000 \times D$ were generated for $d=0.2$, $0.4$, $0.6$, and $0.8$, and 
$D=1$, $101$, $201$, $301$, $401$, and $501$. Since the scikit-learn version of the algorithm
does not exploit sparsity, we assume the method's time is invariate with respect the density of the matrix, but
we gave it the lowest density of each dimensionality to be fair. The results are shown in figure \eqref{fig:benchmark}.

\begin{figure}
\includegraphics[scale=0.3]{benchmark_results.png}
\caption{This plot shows how long the sparse polynomial method took to run on matrices with 1000 rows of various dimensionalities and densities and compared to the scikit-learn PolynomialFeatures class on various dimensionalities.}
\label{fig:benchmark}
\centering
\end{figure}

@@

\begin{center}
    \begin{tabular}{c | c | c} \label{table:results}
 %       \hline
        dataset             & 20newsgroups  & connect-4 \\ \hline
        instances           & 11,314        & 67,557    \\ \hline
        features            & 130,107       & 126       \\ \hline
        polynomial features & 8,463,980,778 & 8,127     \\ \hline
        density             & 0.12          & 0.11      \\ \hline
        dense space (MB)    & Memory Error  & 4191      \\ \hline
        sparse space (MB)   & 5333          & 735       \\ \hline
        dense time (s)      & Memory Error  & 26        \\ \hline
        sparse time (s)     & 109           & 44        \\ \hline
%        \hline
    \end{tabular}
\end{center}

>>>>

\documentclass[11pt,onecolumn]{article}

\usepackage{array}
\usepackage{clrscode3e}
\usepackage{amsmath}
\usepackage{kbordermatrix}
\usepackage{mathtools}
\usepackage{cite}
\usepackage{hyperref}
\usepackage{rotating}
\usepackage{pdfpages}

@@

\section{Introduction}

When performing a modeling task, one is often faced with a non-linearly separable problem,
or is attempting to predict a target that is a non-linear combination of the variables at one's disposal.
There are several general classes of methods for dealing with these situations: kernel methods, 
high variance models, and input transformations.

@@

@@

\begin{sidewaysfigure*}[ht!]
    \centering
    \includegraphics[scale=0.5]{classifier_boundaries.png}
    \caption{Decision boundaries for four classifier types on four different datasets. The darker the color, the more certain the decision.
             The plain dataset is shown in the first column. Each other column is a type of classifier applied to that dataset. The classifier
             types are logistic regression, logistic regression with second degree polynomial features, RBF SVM, and gradient boosting with decision trees.
             The accuracy of each model is shown in the bottom right corner of that model's subplot.}
    \label{fig:boundaries}
\end{sidewaysfigure*}

@@

INSERT JOHN'S PROOF HERE

@@

@@

\subsection{Empirical}
To assess the performance of the algorithm on data of varied density, we compare its speed
to the PolynomialFeatures class of the popular machine learning scikit-learn \cite{scikit-learn}.
Uniformly random matrices of size $1000 \times D$ were generated for $d=0.2$, $0.4$, $0.6$, and $0.8$, and 
$D=1$, $101$, $201$, $301$, $401$, and $501$. Since the scikit-learn version of the algorithm
does not exploit sparsity, we assume the method's time is invariate with respect the density of the matrix, but
we gave it the lowest density of each dimensionality to be fair. The results are shown in figure \eqref{fig:benchmark} on page \pageref{fig:benchmark}.

\begin{figure*}[ht!]
    \includegraphics[scale=0.3]{benchmark_results.png}
    \caption{This plot shows how long the sparse polynomial method took to run on matrices with 1000 rows of various dimensionalities and densities and compared to the scikit-learn PolynomialFeatures class on various dimensionalities.}
    \label{fig:benchmark}
    \centering
\end{figure*}

@@

\begin{center}
    \begin{tabular}{c | c | c} \label{table:results}
 %       \hline
        dataset             & 20newsgroups  & connect-4 \\ \hline
        instances           & 11314        & 67557    \\ \hline
        features            & 130107       & 126       \\ \hline
        polynomial features & 8463980778 & 8,127     \\ \hline
        density             & 0.12          & 0.11      \\ \hline
        dense space (MB)    & Memory Error  & 4191      \\ \hline
        sparse space (MB)   & 5333          & 735       \\ \hline
        dense time (s)      & Memory Error  & 26        \\ \hline
        sparse time (s)     & 109           & 44        \\ \hline
%        \hline
    \end{tabular}
\end{center}
